Summary of the Iberian Bees Database

### 2020-05-22 Summary.Rmd
Create summaries

---
title: "Summary of the Iberian Bees Database"
author: I. Bartomeus, T. Wood, L. Carvalheiro et al.
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

@@

```{r include=FALSE}
data <- read.csv("data/data_clean.csv")
master <- read.csv("data/Iberian_species_masterlist.csv", sep = ";", stringsAsFactors = FALSE)
colnames(data)
colnames(master)
```

The Iberian Bee Database contains `r nrow(data)` entries of bee occurrences over the peninsula and nearby islands. We have at least one record of `r length(unique(data$accepted_name))` species out of the `r nrow(master)` species known to occurr in the area. Here we present a summary of the database:

@@

We have records all over the peninsula, but higer concentrations are marked in red.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#map
library(rgdal)
library(ggplot2)

@@

```{r echo=FALSE}
#barplot of top 10 species
data2 <- subset(data, accepted_name != "Apis mellifera")
default <- par()$mar
par(mar = c(10.1, 4.1, 4.1, 2.1))
barplot(head(sort(table(data2$accepted_name), decreasing = TRUE),20), las = 2, cex.names = 0.7)
par(mar = default)
```

@@

```{r echo=FALSE, warning=FALSE}
hist(data2$Year, main = "Most records are recorded since year 2000", 
     las = 1, xlab = "Year")
#Fix when month is fixed!
hist(data2$Month, main = "Most records are recorded in April",
     las = 1, xlab = "Month", breaks = 12)
```

@@

This is a large collaborative effort lead by I. Bartomeus, L. Carvalheiro and T. Woods, but it would not be possible without the contribution of thousends of researchers. Those are the main contributors (this can be prettier and more data is in the pipeline).

```{r echo=FALSE, message=FALSE, warning=FALSE}
#table of
#Name	
#Contact	
#Number of records as main dataholder	
#Number of records as co-dataholder
#email?? Maually :(
default <- par()$mar
par(mar = c(13.1, 4.1, 4.1, 2.1))
barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
par(mar = default)
#Can be done prettier.
#Ideally, we can merge all authors, split and count individually.
```

>>>>

---
title: "Summary of the Iberian Bees Database"
author: I. Bartomeus, T. Wood, L. Carvalheiro, C. Molina, et al. - 
date: "`r format(Sys.time(), '%d %B, %Y')`"
#output: pdf_document
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

```{r include=FALSE}
data <- read.csv("data/data_clean.csv", stringsAsFactors = FALSE)
master <- read.csv("data/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
colnames(data)
colnames(master)
data[which(!data$accepted_name %in% master$accepted_name),] #OK
master[which(!master$accepted_name %in% data$accepted_name),] 
#we are missing ~100 species, mostly parasites!
length(unique(data$accepted_name))
nrow(master)
```

The Iberian Bee Database contains `r nrow(data)` entries of bee occurrences over the peninsula and nearby islands. We have at least one record of `r length(unique(data$accepted_name))` species out of the `r nrow(master)` species known to occurr in the area.

We retained only valid, updated species names. All records have accompaining metadata on where and when it was collected, and on who collected, identified and curated the data. We are still reciving and cleaning data, but we aim to publish the open database after summer.

Here we present a summary of the database:

@@

We have records all over the peninsula, but higer concentrations are marked in red.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#map
library(rgdal)
library(ggplot2)

@@

```{r echo=FALSE}
#barplot of top 10 species
data2 <- subset(data, accepted_name != "Apis mellifera")
default <- par()$mar
par(mar = c(10.1, 4.1, 4.1, 2.1))
barplot(head(sort(table(data2$accepted_name), decreasing = TRUE),20), las = 2, cex.names = 0.7)
#barplot(tail(sort(table(data2$accepted_name), decreasing = TRUE),20), las = 2, cex.names = 0.7)
par(mar = default)
```

@@

```{r echo=FALSE, warning=FALSE}
hist(data2$Year, main = "Most records are recorded since year 2000", 
     las = 1, xlab = "Year")
```

```{r echo=FALSE, warning=FALSE}
hist(data2$Month, main = "Most records are recorded in April",
     las = 1, xlab = "Month", breaks = 12)
```

@@

This is a large collaborative effort lead by I. Bartomeus, L. Carvalheiro and T. Woods, but it would not be possible without the contribution of thousends of researchers. Those are the main contributors, but more contributors will be joining soon!

```{r echo=FALSE, message=FALSE, warning=FALSE}
#table of
#Name	
#Contact	
#Number of records as main dataholder	
#Number of records as co-dataholder
#email?? Maually :(
#default <- par()$mar
#par(mar = c(13.1, 4.1, 4.1, 2.1))
#barplot(head(sort(table(data2$Authors.to.give.credit), decreasing = TRUE),20), las = 2, cex.names = 0.7)
#par(mar = default)
#Can be done prettier.
#Ideally, we can merge all authors, split and count individually.
library(stringi)
authors <- data$Authors.to.give.credit[which(!is.na(data$Authors.to.give.credit))]
authors[which(authors == "Ornosa, C., Martínez, M. D.")] <- "C. Ornosa, M.D. Martínez"
authors[which(authors == "C. Ornosa & F. J. Ortiz-Sánchez")] <- "C. Ornosa, F. J. Ortiz-Sánchez"
authors[which(authors == "C. Ornosa & M.D. Martínez")] <- "C. Ornosa, M.D. Martínez"
authors[which(authors == "Ortiz-Sanchez, F.J.")] <- "F.J. Ortiz-Sanchez"
authors[which(authors == "Heras, C., Gayubo, S. F.")] <- "C. Heras, S.F. Gayubo"
authors[which(authors == "Ornosa, C.")] <- "C. Ornosa"
authors[which(authors == "Herrera, J.")] <- "Javier Herrera"
authors[which(authors == "Francisco Javier Ortiz-Sánchez, Pedro Coello & Manuel Baena")] <- "Francisco Javier Ortiz-Sánchez, Pedro Coello, Manuel Baena"
authors[which(authors == "González, J. A., Torres, F., Gayubo, S. F.")] <- "J.A. González, , F. Torres, S.F. Gayubo"
authors[which(authors == "Dardón, M. J., Torres, F., Asís, J. D.")] <- "M.J. Dardón, F. Torres, J.D. Asís"
authors[which(authors == "Pérez-Iñigo Mora, C.")] <- "C. Pérez-Iñigo Mora"
authors[which(authors == "Pablo Vargas, Concepción Ornosa, Francisco Javier Ortiz-Sánchez & Juan Arroyo")] <- "Pablo Vargas, Concepción Ornosa, Francisco Javier Ortiz-Sánchez, Juan Arroyo"
authors[which(authors == "C. Heras y S.F. Gayubo")] <- "C. Heras, S.F. Gayubo"
authors[which(authors == "J. L. BLANCO-PASTOR, C. ORNOSA, D. ROMERO, I. M. LIBERAL, J. M. GÓMEZ & P. VARGAS")] <- "J. L. Blanco-Pastor, C. Ornosa, D. Romero, I.M. Liberal, J.M. Gómez, P. Vargas"
authors[which(authors == "Concepción Ornosa, Félix Torres & Pilar de la Rúa")] <- "Concepción Ornosa, Félix Torres, Pilar de la Rúa"
authors[which(authors == "Castro, L.")] <- "L. Castro"
authors[which(authors == "Ortiz-Sanchez, F. J., Ornosa, C.")] <- "F.J. Ortiz-Sanchez, C. Ornosa"
authors[which(authors == "Banaszak, J., Ortiz-Sanchez, F. J.")] <- "J. Banaszak, F.J. Ortiz-Sanchez"
authors[which(authors == "C. Ornosa, F. Torres & F. J. Ortiz-Sánchez")] <- "C. Ornosa, F. Torres, F. J. Ortiz-Sánchez"

author_pool <- stri_paste(authors, collapse = ",")
author_list <- unlist(strsplit(author_pool, ","))
author_list <- trimws(author_list)
#sort(unique(author_list))[-1] #79 contributors!
#write.csv(sort(unique(author_list))[-1], "data/authors.csv")

auth <- c(
"Alejandro Núñez"
,"Amparo Lázaro"
,"Ana Montero-Castaño"
,"Anna Traveset"
,"Blanca Arroyo-Correa"
,"C. Heras"
,"Curro Molina"
,"C. O’Connor"
,"Concepción Ornosa"
,"C. Pérez-Iñigo Mora"
,"Carlos Lara-Romero"
,"T. Wood"
,"Concepción Ornosa"
,"Concepción Ornosa"
,"D. Paredes"
,"D. Romero"
,"David Alomar"
,"F. Javier Ortiz-Sánchez"
,"F. Javier Ortiz-Sánchez"
,"F. López-Núñez"
,"F. Perfectti"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"Félix Torres"
,"F. Javier Ortiz-Sánchez"
,"H. Gaspar"
,"Ignasi Bartomeus"
,"Isabel María Liberal"
,"Isabel María Liberal"
,"J. Ascher"
,"J. Banaszak"
,"J. Bosch"
,"J. Dusmet"
,"J. Gonzalez-Varo"
,"J. L. Blanco-Pastor"
,"J. Loureiro"
,"J. R. Obeso"
,"J.A. González"
,"J. Bosch"
,"J.D. Alche"
,"J.D. Asís"
,"J.M. Gómez"
,"Javier Herrera"
,"Javier Morente-López"
,"Joana Costa"
,"José Iriondo"
,"Jose Luis Blanco-Pastor"
,"Juan Arroyo"
,"L. Castro"
,"Luis Navarro"
,"M. Campos"
,"M. Miñarro"
,"M. Porcel"
,"M.D. Martínez"
,"M.J. Dardón"
,"M.L. Fernández"
,"Montserrat Vilà"
,"Manuel Baena"
,"Marina Alonso Solano"
,"Mario Fernández-Mazuecos"
,"Carlos Martínez-Núñez"
,"Miguel A. González-Estévez"
,"Montserrat Vilà"
,"Pablo Vargas"
,"Pablo Vargas"
,"Pedro Coello"
,"Pilar de la Rúa"
,"R. Alcalá"
,"R. Carvalho"
,"R. Heleno"
,"P.J. Rey"
,"S. Castro"
,"S. Castro"
,"S.F. Gayubo"
,"Victoria Ferrero"
,"Volker Haeseler")

unique(auth)
```

The list of co-authors is not exhaustive, and we did not carefully checked contributor names yet. Contact nacho.bartomeus@gmail.com if you want to contribute.
